Rapid-response Bayesian sky localization for electromagnetic follow-up of gravitational-wave candidates

### 2013-03-08 ms.tex
Removed double-spaces between sentences
Removed duplicate copies of some text

A \ac{CBC} source is specified by a vector of extrinsic parameters describing its position and orientation and intrinsic parameters describing the physical properties of the binary components:
%
\begin{equation}
    \boldsymbol\theta = \begin{array}{l@{}l}
            \left[
            \begin{array}{c}
                \alpha \\
                \delta \\
                D_\mathrm{L} \\
                t_\oplus \\
                \iota \\
                \psi \\
                \phi_c \\
                \hdashline[1pt/1pt]
                m_1 \\
                m_2 \\
                \mathbf S_1 \\
                \mathbf S_2
            \end{array}
            \right] &
            \begin{array}{l}
                \left.
                \begin{array}{p{4cm}}
                    right ascension \\
                    declination \\
                    luminosity distance \\
                    arrival time at geocenter \\
                    inclination angle \\
                    polarization angle \\
                    coalescence phase \\
                    \hdashline[1pt/1pt]
                \end{array}
                \quad \right\} \textnormal{extrinsic} \\
                \left.
                \begin{array}{p{4cm}}
                    first component's mass \\
                    second component's mass \\
                    first component's spin \\
                    second component's spin
                \end{array}
                \quad \right\} \textnormal{intrinsic}
            \end{array}
        \end{array}
\end{equation}
%
\marginpar{Citations needed. Do we want to mention these omitted parameters?}%
%
Binary eccentricity is omitted as an intrinisc parameter because \acp{GW} circularize the orbit long before the inspiral enters \ac{LIGO}'s frequency range of $\sim$10\nobreakdashes--1000~kHz. Tidal deformabilities of the components are omitted because tidal effects have a negligible effect in the inspiral phase.

In this work we restrict ourselves to nonspinning sources, where $\mathbf{S}_1=\mathbf{S}_2=0$. Removing nuisance parameters requires either replacing them with their measured values or marginalizing over them. The chirp mass is typically well measured by the detection pipeline, so we use the maximum likelihood value, $\hat{\mathcal{M}}$, for it. We marginalize over the other variables using a uniform prior, i.e.
\begin{equation}
p(\alpha,\delta,D_L|d) = \int\, d\tau_\oplus p_U(\tau_\oplus) d\phi_c p_U(\phi_c) d\phi p_U(\psi)  d\cos(\iota) p_U(\cos(\iota)) p(\alpha,\delta,D_L,\tau_\oplus,\phi_c,\psi,\cos(\iota)|d),
\end{equation}
where $p_U(x)$ denotes a uniform distribution in $x$.

@@

In the full signal model the paramters, $\boldsymbol\theta$, are given by a 15-dimensional vector of quantities. Of those, the quantities of interest are the right ascenion and declination, $(\alpha,\delta)$, and luminosity distance, $D_L$. The remaining quantites, or {\it nuisance parameters} for the full signal model are:
\begin{itemize}
\item The chirp mass. If $m_1$ and $m_2$ are the masses of the individual components then $\mathcal{M}=(m_1m_2)^{3/5}/(m_1+m_2)^{1/5}$.
\item The symmetric mass ratio, $\eta=(m_1m_2)/(m_1+m_2)^2$.
\item The 3-dimensional spin vectors of each components, $\mathbf{S}_1,\mathbf{S}_2$.
\item An arrival time, $\tau_\oplus$, taken here to be the time of arrival at the geocenter.
\item The inclination angle, $\iota$.
\item The polarization angle, $\psi$.
\item The coalescence phase, $\phi_c$.
\end{itemize}
In this work we restrict ourselves to non-spinng sources, where $\mathbf{S}_1=\mathbf{S}_2=0$. Removing nuisance parameters requires either replacing them with their measured values or marginalizing over them. The chirp mass is typically well measured by the detection pipeline, so we use the maximum likelihood value, $\hat{\mathcal{M}}$, for it. We marginalize over the other variables using a uniform prior, i.e.
\begin{equation}
p(\alpha,\delta,D_L|d) = \int\, d\tau_\oplus p_U(\tau_\oplus) d\phi_c p_U(\phi_c) d\phi p_U(\psi)  d\cos(\iota) p_U(\cos(\iota)) p(\alpha,\delta,D_L,\tau_\oplus,\phi_c,\psi,\cos(\iota)|d),
\end{equation}
where $p_U(x)$ denotes a uniform distribution in $x$.

>>>>

A \ac{CBC} source is specified by a vector of extrinsic parameters describing its position and orientation and intrinsic parameters describing the physical properties of the binary components:
%
\begin{equation}
    \boldsymbol\theta = \begin{array}{l@{}l}
            \left[
            \begin{array}{c}
                \alpha \\
                \delta \\
                D_\mathrm{L} \\
                t_\oplus \\
                \iota \\
                \psi \\
                \phi_c \\
                \hdashline[1pt/1pt]
                m_1 \\
                m_2 \\
                \mathbf S_1 \\
                \mathbf S_2
            \end{array}
            \right] &
            \begin{array}{l}
                \left.
                \begin{array}{p{4cm}}
                    right ascension \\
                    declination \\
                    luminosity distance \\
                    arrival time at geocenter \\
                    inclination angle \\
                    polarization angle \\
                    coalescence phase \\
                    \hdashline[1pt/1pt]
                \end{array}
                \quad \right\} \textnormal{extrinsic} \\
                \left.
                \begin{array}{p{4cm}}
                    first component's mass \\
                    second component's mass \\
                    first component's spin \\
                    second component's spin
                \end{array}
                \quad \right\} \textnormal{intrinsic}
            \end{array}
        \end{array}
\end{equation}
%
\marginpar{Citations needed. Do we want to mention these omitted parameters?}%
%
Binary eccentricity is omitted as an intrinisc parameter because \acp{GW} circularize the orbit long before the inspiral enters \ac{LIGO}'s frequency range of $\sim$10\nobreakdashes--1000~kHz. Tidal deformabilities of the components are omitted because tidal effects have a negligible effect in the inspiral phase. In \ac{GW} detection efforts, especially those focused on \ac{NSNS} systems, the component spins $\mathbf{S}_1$ and $\mathbf{S}_2$ are often assumed to be aligned with the system's total angular momentum and condensed to a single scalar parameter $\chi$, or even neglected entirely, $\mathbf{S}_1 = \mathbf{S}_2 = 0$. Without loss of generality, in this work we will consider only zero spin.

Removing nuisance parameters requires either replacing them with their measured values or marginalizing over them. The chirp mass is typically well measured by the detection pipeline, so we use the maximum likelihood value, $\hat{\mathcal{M}}$, for it. We marginalize over the other variables using a uniform prior, i.e.
\begin{equation}
p(\alpha,\delta,D_L|d) = \int\, d\tau_\oplus p_U(\tau_\oplus) d\phi_c p_U(\phi_c) d\phi p_U(\psi)  d\cos(\iota) p_U(\cos(\iota)) p(\alpha,\delta,D_L,\tau_\oplus,\phi_c,\psi,\cos(\iota)|d),
\end{equation}
where $p_U(x)$ denotes a uniform distribution in $x$.

@@
